### 2020-10-08 paper/article.Rmd

---
title:  |
        | Long Article Title:
        | A Very Long Subtitle
author: 
  - Danilo Freire[^freire]
  - Umberto Mignozzetti[^mignozzetti]
date: \today
thanks: "The authors thank Professors John and Jane Doe for their helpful comments."
abstract: "This is a personal article template for RMarkdown. It uses Linux Libertine and Inconsolata fonts, British English, double-spaced text and numbered sections. The colours are personalised too: URLs are in dark blue, internal links are in mahogany and the table of contents is in black. The template also has pre-formatted entries for the abstract, keywords and JEL classification codes. You need to install `R` (and the `rmarkdown` package), `pandoc` and a `TeX` system to compile the pdf. In case you don't want to download large `TeX` binaries, I recommend you to install the great `TinyTeX R` package. A quick note: if you want to use colons in the abstract or the `thanks` entry, please don't forget to put the whole text in quotes. Otherwise, YAML might not parse it correctly. I hope you find this template useful and feel free to send me a message if you have any suggestions."
abstractspacing: double
keywords: R Markdown; pandoc; template
jelcodes: A00, B11, C22
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# If you need to install any package while knitting the document
r <- getOption("repos")
r["CRAN"] <- "https://cran.rstudio.com/"
options(repos = r)
if (!require("kableExtra")) {
    install.packages("kableExtra")
}
```

[^freire]: Independent Researcher,
[`danilofreire@gmail.com`](mailto:danilofreire@gmail.com),
<http://danilofreire.github.io>.

[^mignozzetti]: Visiting Assistant Professor, Quantitative Theory and Methods
Department, Emory University, 532 North Kilgo Circle, 4^th^ Floor, Atlanta, GA
30322, USA,
[`umberto.mignozzetti@emory.edu`](mailto:umberto.mignozzetti@emory.edu),
<http://umbertomig.com>.

\newpage

# Introduction
\label{sec:intro}

The *Aedes aegypti* mosquito species is the major vector of several arboviral
diseases worldwide, including dengue, yellow fever, Zika, and chikungunya
[@burt2012chikungunya; @jansen2010dengue; @leta2018global;
@thangamani2016vertical]. Over the past 30 years, arboviral incidence has
increased exponentially, with estimates suggesting that dengue is endemic in
around 100 countries, causing 400 million infections and 22 thousands deaths
each year [@bhatt2013global; @cdc2020dengue; @musso2018unexpected]. Since there
are no safe vaccines for most *A. aegypti*-transmitted diseases
[@sridhar2018effect; @world2017dengue], we have seen a number of recent dengue
and Zika outbreaks in Asia, the Caribbean, South America, and the United States
[@carlson2016ecological; @leta2018global]. Chikungunya virus, which was endemic
only to Africa, has already been reported in every region of the world
[@grandadam2011chikungunya; @kraemer2015global; @mayer2017emergence].

Mosquito control is the primary strategy to prevent *A. aegypti*-borne
infections [@beatty2010best; @eisen2009proactive; @ooi2006dengue]. The World
Health Organization recommends the elimination of containers and the use of
insecticide to reduce vector breeding, specially in settings where human-vector
contacts are frequent, such as schools, hospitals, and workplaces
[@WHO2012global; @world2020denguecontrol]. However, cleaning breeding sites and
insecticide spraying are labour-intensive processes and are often carried out
inconsistently [@paz2015dengue; @siddiqui2016use; @WHO2012global]. In most
endemic countries, local governments are the main responsible for providing
vector control, thus increasing the productivity of civil servants is essential
to the effectiveness of anti-mosquito programmes worldwide.

Financial incentives are crucial determinants of worker performance in public
and private organisations [@burgess2003role]. While studies indicate that
reward schemes may improve the productivity of health care providers
[@basinga2011effect; @flodgren2011overview; @ulrich2005does], it is unclear how
they affect the motivation of workers fighting *A. aegypti*. Moreover, there is
no evidence about the impact of financial incentives on patient-level arboviral
outcomes, such as virus-related hospitalisations or disease incidence
[@flodgren2011overview].

To explore how monetary incentives affect health worker motivation and patient
outcomes, we ran a randomised field experiment in the city of Rio Verde,
Brazil, which has experienced a peak in dengue infections in the last years. We
hired and trained teams of health care workers to perform simplified procedures
designed to exterminate *A. aegypti* breeding sites. Our experiment consisted
of two treatment arms plus a control condition. The control group received a
fixed compensation for completing their task, and we offered individual or
collective bonuses to those allocated in the two treatment groups. We then
measured the impact of the compensations schemes on the number of breeding
sites discovered, larvae exterminated, and city-level dengue incidence 30, 60,
and 120 days after our intervention.

We find that monetary rewards increase the number of cleaned breeding sites in
both treatment conditions (individual and team bonuses) and collective
financial incentives also improve larvae extermination. Our results show that
the control group has a high incidence of houses visited in less than two
minutes apart from each other, which suggests that workers may cheat in the
absence of financial incentives. When we combine both treatments, we also find
that the intervention decreases dengue incidence in 10.3%, but the results are
not robust. In summary, we provide experimental evidence that performance
bonuses increase productivity in public service delivery and that small
financial incentives to health workers may improve patient outcomes and reduce
hospitalisation costs in *Aedes*-endemic countries.

# Methods
\label{sec:methods}

We ran a randomised field experiment at the individual level in the municipality
of Rio Verde, Brazil, which had experienced a major dengue outbreak one year
prior to our experiment (2017). We used Facebook advertisements to recruit
potential participants. The advertisements directed applicants to a Google Forms
web page that requested their contact information, and we sent a pre-treatment
demographic questionnaire to every applicant who provided a valid email address.
At this stage, we excluded all individuals under 18 years old. We invited the
remaining applicants to our training session, which we ran on the 4^th^ of May,
2018. The meeting consisted of a lecture on *A. aegypti* vector biology and
practical instructions on how to find and exterminate mosquito breeding sites.
Those who did not attend the session were also excluded from the study.

Table \ref{tab:covbal} displays the attrition tests, the covariate balance
tests, and the territorial assignment before and after applying propensity
score matching [@dehejia2002propensity; @rosenbaum1985constructing] on census
tract characteristics. We observe no significant differences regarding
attrition rates, reinforcing that the treatment was correctly randomised.
Participants' demographics are also similar across all groups, and the matching
technique successfully balanced the territorial assignment.

\begin{table}[ht] \centering 
\small
  \caption{Covariate Balance on Individual Characteristics} 
  \label{tab:covbal} 
\begin{tabular}{@{\extracolsep{5pt}} lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Variable & Attrition & Control & Individual & Collective & F-Statistic & P-Value \\
\hline \\[-1.8ex] 
& & \multicolumn{5}{l}{Panel A: Attrition Pattern Test} \\
\hline \\[-1.8ex]
Age &  26.17 & 25.04 & 26.07 & 26.00 & 0.461 & 0.710 \\ 
Female &   0.72 &  0.79 &  0.79 &  0.74 & 0.805 & 0.492 \\ 
Has Vehicle &   0.35 &  0.27 &  0.37 &  0.37 & 0.651 & 0.582 \\ 
N & 341 & 73 & 71 & 73 &  &  \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\begin{tabular}{@{\extracolsep{5pt}} lccccc} 
Variable & Control & Individual & Collective & F-Statistic & P-Value \\
\hline \\[-1.8ex] 
& \multicolumn{5}{l}{Panel B: Covariate Balance on Individual Characteristics} \\
\hline \\[-1.8ex]
Altruist &  0.63 &  0.68 &  0.62 & 0.239 & 0.788 \\ 
Age & 25.01 & 26.00 & 26.90 & 0.754 & 0.472 \\ 
Religiosity &  0.53 &  0.38 &  0.45 & 1.487 & 0.228 \\ 
Pol. Engagement &  0.69 &  0.65 &  0.68 & 0.163 & 0.849 \\ 
Soc. Engagement &  0.59 &  0.62 &  0.67 & 0.454 & 0.636 \\ 
Above 2 Min. Wage &  0.34 &  0.35 &  0.29 & 0.336 & 0.715 \\ 
FB Popularity &  0.53 &  0.56 &  0.58 & 0.175 & 0.840 \\ 
\hline \\[-1.8ex] 
& \multicolumn{5}{l}{Panel C: Unmatched Covariate Balance on Territorial Assignment} \\
\hline \\[-1.8ex] 
Number of Houses & 276.41 & 243.58 & 255.20 &  2.368 & 0.098 \\ 
Number of Households & 885.67 & 750.34 & 785.18 &  3.765 & 0.026 \\ 
Avg. Household Size &   3.21 &   3.07 &   3.07 &  6.566 & 0.002 \\ 
Log Avg. Income &   6.47 &   6.86 &   6.85 &  9.213 & 0.000 \\ 
Cases Before Treatment &  13.78 &   5.94 &   4.66 & 15.638 & 0.000 \\ 
\hline \\[-1.8ex] 
& \multicolumn{5}{l}{Panel D: Matched Covariate Balance on Territorial Assignment} \\
\hline \\[-1.8ex] 
Number of Houses & 276.41 & 282.04 & 281.19 & 0.068 & 0.935 \\ 
Number of Households & 885.67 & 903.04 & 900.39 & 0.060 & 0.942 \\ 
Avg. Household Size &   3.21 &   3.21 &   3.20 & 0.008 & 0.992 \\ 
Log Avg. Income &   6.47 &   6.45 &   6.45 & 0.084 & 0.920 \\ 
Cases Before Treatment &  13.78 &  10.24 &  10.77 & 1.443 & 0.240 \\ 
\hline
Pairs & 34 & 34 & 35 &  &  \\ 
N & 68 & 68 & 70 &  &  \\
\hline\hline \\[-1.8ex] 
\end{tabular}
\end{table}

\normalsize

The intervention took place on the 5^th^ of May, 2018. We randomly assigned
subjects to three groups. To avoid spillovers, we directed each group to
separate headquarters before the intervention began. Participants worked in
pairs and were aware of other groups, but they did not know about the different
treatment conditions. In the control group ($n = 68$), we told participants
that they would receive a flat payment regardless of their performance (BRL 110
/ USD 20 per subject). In the individual treatment group ($n = 68$), we
informed subjects that we would rank individual performances, and that we would
double the compensation for all of those who performed above the sample median
(BRL 220 / USD 50 per subject). In the collective treatment group ($n = 70$),
we explained that we would rank the pair performance and double the payment for
teams who presented results better than the sample median (BRL 220 / USD 50 per
subject). The teams received a leaflet with a map of the area they should
cover. The route comprised of two to three blocks and contained around 120
houses.

A typical visit to a household consisted of the following steps. First,
participants rang the doorbell in a house and explained that the municipality
had experienced a dengue fever outbreak. They instructed dwellers on how to
lower dengue incidence and handed them a leaflet with information about dengue
prevention practices. Our subjects then requested whether they could inspect
the person's yard. If granted permission, they entered the household and
searched for clean breeding sites, such as pots filled with clear water, and
recipients containing larvae. Upon exterminating a breeding site (pouring out
the accumulated water, for example), subjects had to report their task by
taking a photo on their cellphones. When they discovered larvae, they had to
record a video showing the larvae before exterminating it. At the end of their
visit, they had to take a picture of the household to account for their
presence in the location.

The primary outcome in this study is field productivity of the health care
workers. We measure worker productivity by coding cellphone data collected
during the intervention into four indicators: 1) the number of houses visited;
2) the number of houses visited within less than two minutes; 3) the number of
breeding sites removed and cleaned; 4) the number of exterminated larvae. We
employed ordinary least squares with robust standard errors [@imbens2016robust].
We estimated the statistical power of the study assuming a treatment effect of
0.15 standard deviations.

The second outcome of interest is dengue incidence in the municipality of Rio
Verde. The Rio Verde Mayor's Office provided us with its annual health care
report, which contains the dates and personal information of patients who had
dengue fever and received hospital treatment from January to September 2018. We
georeferenced the home addresses of the patients and compared the incidence map
with the distribution of our intervention. We use a differences-in-differences
estimator to assess the impact of our experiment [@angrist2008mostly;
@card1994minimum].

We pre-registered the experiment on the Evidence in Governance and Policy
(EGAP) website (<https://osf.io/6q8vu/>) and received IRB approval from New
York University (IRB-FY2017-17) and Fundação Getulio Vargas (IRB-01/2017). The
funders of this study had no influence over the formulation, organisation,
analysis, and interpretation of the outcomes. Health care workers gave written
consent upon filling the Qualtrics questionnaire. During the intervention, the
occupants of households inspected for breeding sites and larvae gave oral
consent. For more information on the experimental design, sample
characteristics, exclusion criteria, and estimation methods, please refer to
the Supplementary Material.

# Results
\label{sec:results}

We find that providing financial incentives to health care worker significantly
increase their field productivity. Both our individual bonus and our collective
bonus treatment groups find more *A. aegypti* breeding sites and exterminate
more mosquito larvae than participants in the control condition, which did not
receive performance compensations. Our results also indicate that the
intervention decreases city-wide levels of dengue incidence, but the
coefficients only reach statistical significance when we combine the two
treatments together and contrast them with the control group. Overall, the
findings suggest that monetary payments are an effective method of improving
health worker performance, yet more research is needed to assess whether higher
labour productivity leads to fewer dengue-related hospitalisations.

## Field Productivity
\label{sub:productivity}

We have four indicators to measure health worker productivity. The first is the
number of houses visited during the intervention. The second indicator is the
number of houses in which the visits lasted for less than two minutes. The
third dependent variable is the number of *A. aegypti* breeding sites removed
and cleaned. Our fourth variable indicates the number of mosquito larvae
exterminated. Table \ref{tab:productivity} summarises the experimental results.

\begin{table}[ht] 
  \centering 
  \small
  \caption{Field Productivity} 
  \label{tab:productivity} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent Variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & Houses & Houses & Breeding & Larvae \\ 
& Visited & Visited ($<$ 2min.) & Sites & Exterminated \\ 
\hline \\[-1.8ex] 
 Individual Bonus & $-$9.879$^{**}$ & $-$8.044$^{**}$ & 25.118$^{***}$ & 0.022 \\ 
  & (4.906) & (4.018) & (5.251) & (0.058) \\ 
  & & & & \\ 
 Collective Bonus & $-$7.554$^{*}$ & $-$6.479$^{*}$ & 21.512$^{***}$ & 0.163$^{**}$ \\ 
  & (4.428) & (3.571) & (4.049) & (0.076) \\ 
  & & & & \\ 
\hline \\[-1.8ex] 
Observations & 197 & 197 & 197 & 197 \\ 
Residual SE & 27.301 & 22.153 & 30.545 & 0.425 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{Robust standard errors in parentheses. $^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

Subjects who received monetary incentives, either alone or in pairs, visited
fewer houses than those who did not earn financial rewards. While the result
may seem counter-intuitive, it actually attests the effectiveness of the
treatment. This is because we offered incentives for subjects to exterminate
larvae and find breeding sites, but not to visit a greater number of
households. As a result, participants in the two treatment groups did a
thorough job in each household, which explains the lower number of visits
during the allotted time. We find that subjects in the individual and in the
collective bonus groups visited $-9.879$ (CI: [XXX, XXX], $p-$value $= 0.XXX$)
and $-7.554$ (CI: [XXX, XXX], $p-$value $= 0.XXX$) households, respectively.

Our second indicator provides further evidence in favour of our argument. We
see that the two groups which earned performance bonuses spent more time in the
households, as their number of houses visited for less than two minutes is
significantly lower than those of the control condition. On average,
participants in the individual bonus treatment arm visited $-8.044$ (CI: [XXX,
XXX], $p-$value $= 0.XXX$) fewer houses than the control, and those in the
collective bonus group visited $-6.479$ (CI: [XXX, XXX], $p-$value $= 0.XXX$).

Incentivised teams find and clean more breeding sites than non-incentivised
ones. Individually incentivised health workers find 25.118 (CI: [XXX, XXX],
$p-$value $= 0.XXX$) more breeding sites than the non-incentivised control
group. The effect for those in the collective bonus group is 21.512 (CI: [XXX,
XXX], $p-$value $= 0.XXX$), slightly smaller than for the first treatment.

Collective incentivized teams are more effective in finding larvae. In our
study, collectively incentivized subjects have 16.3\% more chance to find
larvae than non incentivized workers. Finding larvae is harder, and the
detection chances profit from teamwork. If teams are cooperating, searching the
house becomes more natural, and they can coordinate to do a more thorough job.
Additionally, the detection tends to be excludable: one individual finding
larvae makes the chances of keep finding larvae in the household considerably
smaller. Collective bonus is not excludable within the teams, making them
cooperate more to get the reward.

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{5pt}