### 2020-10-09 paper/article.Rmd
WIP

---
title:  |
        | Long Article Title:
        | A Very Long Subtitle
author: 
  - Danilo Freire[^freire]
  - Umberto Mignozzetti[^mignozzetti]
date: \today
thanks: "The authors thank Professors John and Jane Doe for their helpful comments."
abstract: "This is a personal article template for RMarkdown. It uses Linux Libertine and Inconsolata fonts, British English, double-spaced text and numbered sections. The colours are personalised too: URLs are in dark blue, internal links are in mahogany and the table of contents is in black. The template also has pre-formatted entries for the abstract, keywords and JEL classification codes. You need to install `R` (and the `rmarkdown` package), `pandoc` and a `TeX` system to compile the pdf. In case you don't want to download large `TeX` binaries, I recommend you to install the great `TinyTeX R` package. A quick note: if you want to use colons in the abstract or the `thanks` entry, please don't forget to put the whole text in quotes. Otherwise, YAML might not parse it correctly. I hope you find this template useful and feel free to send me a message if you have any suggestions."
abstractspacing: double
keywords: R Markdown; pandoc; template
jelcodes: A00, B11, C22
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: template.latex
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# If you need to install any package while knitting the document
r <- getOption("repos")
r["CRAN"] <- "https://cran.rstudio.com/"
options(repos = r)
if (!require("kableExtra")) {
    install.packages("kableExtra")
}
```

[^freire]: Independent Researcher,
[`danilofreire@gmail.com`](mailto:danilofreire@gmail.com),
<http://danilofreire.github.io>.

[^mignozzetti]: Visiting Assistant Professor, Quantitative Theory and Methods
Department, Emory University, 532 North Kilgo Circle, 4^th^ Floor, Atlanta, GA
30322, USA,
[`umberto.mignozzetti@emory.edu`](mailto:umberto.mignozzetti@emory.edu),
<http://umbertomig.com>.

\newpage

# Introduction
\label{sec:intro}

The *Aedes aegypti* mosquito is the major worldwide vector of several arboviral
diseases, including dengue, yellow fever, Zika, and chikungunya
[@burt2012chikungunya; @jansen2010dengue; @leta2018global;
@thangamani2016vertical]. Over the past 30 years, arboviral incidence has
increased exponentially, with estimates suggesting that dengue is endemic in
around 100 countries and causes 400 million infections and 22 thousands deaths
each year [@bhatt2013global; @cdc2020dengue]. Since there are no safe vaccines
for most *Aedes*-transmitted diseases [@sridhar2018effect;
@world2017dengue], we have seen a number of dengue and Zika outbreaks in Asia,
the Caribbean, South America, and the United States [@carlson2016ecological;
@leta2018global]. Chikungunya virus, which was endemic only to Africa, has
already been reported in every region of the world [@grandadam2011chikungunya;
@kraemer2015global; @mayer2017emergence].

Mosquito control is the primary strategy to prevent *Aedes*-borne infections
[@beatty2010best; @eisen2009proactive]. The World Health Organization
recommends the elimination of containers and the use of insecticide to reduce
vector breeding, specially in settings where human-vector contacts are
frequent, such as schools, hospitals, and workplaces [@WHO2012global;
@world2020denguecontrol]. However, breeding site cleaning and insecticide
spraying are labour-intensive processes, and they are often carried out
inconsistently in developing countries [@paz2015dengue; @siddiqui2016use;
@WHO2012global]. In this regard, improving the performance of those responsible
for mosquito control is essential to the effectiveness of anti-*Aedes*
programmes.

Research shows that financial incentives may increase the productivity of
health care workers [@clemens2014physicians; @muthmainnah2017improving;
@ulrich2005does]. However, systematic reviews indicate that the available
evidence has serious methodological shortcomings, such as selection bias and
inadequate experimental blinding, what limits the generalisability of current
findings [@bassani2013financial; @flodgren2011overview; @scott2011effect].
Moreover, it remains unclear which rewarding schemes have a stronger effect on
health care worker performance [@fainman2020design], and whether monetary
incentives have a sustained effect on patient outcomes [@jha2012long;
@langdown2014use].

In this paper, we explore how financial incentives affect the productivity of
health care personnel involved in *A. aegypti* control and measure how monetary
rewards impact dengue hospitalisations. We run a randomised field experiment in
the city of Rio Verde, Brazil, which has experienced a peak in dengue
infections in the last years. We hired and trained teams of health care workers
to perform simplified procedures designed to exterminate *A. aegypti* breeding
sites. Our experiment consisted of two treatment arms plus a control condition.
The control group received a fixed compensation for completing their tasks, and
we offered individual or collective bonuses to those allocated in the two
treatment groups. We then measured how the different compensations schemes
affect the number of breeding sites discovered, larvae exterminated, and
city-level dengue incidence.

We find that monetary rewards increase the number of cleaned breeding sites in
both treatment conditions (individual and team bonuses), and that collective
financial incentives also improve larvae extermination. Our results show that
the control group has a high number of houses visited in less than two minutes
apart from each other, which suggests that workers may cheat in the absence of
financial incentives. When we combine both treatments, we also find that the
intervention decreases dengue incidence in 10.3%, but the results are not
robust. In summary, we provide experimental evidence that performance bonuses
enhance public service delivery and that small financial incentives may improve
patient outcomes and reduce hospitalisations in *Aedes*-endemic countries.

# Methods
\label{sec:methods}

We conducted a randomised field trial in Rio Verde, a municipality in the
Brazilian Midwestern state of Goiás. Rio Verde had 229,000 inhabitants and 1,156
confirmed dengue cases, one of highest counts in the state [@goias2020]. We used
Facebook advertisements to recruit potential participants. The advertisements
directed applicants to Google Forms, where we asked for their contact
information. We sent a pre-treatment demographic survey to every applicant who
provided a valid email address. At this stage, we excluded all individuals under
18 years old. We invited the remaining applicants to our training session, which
we ran on the 4^th^ of May, 2018. The meeting consisted of a lecture on *A.
aegypti* vector biology and practical instructions on how to find and
exterminate mosquito breeding sites. Those who did not attend the session were
also excluded from the study.

Table \ref{tab:covbal} displays the attrition tests, the covariate balance
tests, and the territorial assignment before and after applying propensity score
matching [@rosenbaum1985constructing] on census tract characteristics. We
observe no significant differences regarding attrition rates, indicating that
the treatment was correctly randomised. Participants' demographics are also
similar across all groups, and the matching technique successfully balanced the
territorial assignment.

\begin{table}[ht] \centering 
\small
  \caption{Covariate Balance on Individual Characteristics} 
  \label{tab:covbal} 
\begin{tabular}{@{\extracolsep{5pt}} lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Variable & Attrition & Control & Individual & Collective & F-Statistic & P-Value \\
\hline \\[-1.8ex] 
& & \multicolumn{5}{l}{Panel A: Attrition Pattern Test} \\
\hline \\[-1.8ex]
Age &  26.17 & 25.04 & 26.07 & 26.00 & 0.461 & 0.710 \\ 
Female &   0.72 &  0.79 &  0.79 &  0.74 & 0.805 & 0.492 \\ 
Has Vehicle &   0.35 &  0.27 &  0.37 &  0.37 & 0.651 & 0.582 \\ 
N & 341 & 73 & 71 & 73 &  &  \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\begin{tabular}{@{\extracolsep{5pt}} lccccc} 
Variable & Control & Individual & Collective & F-Statistic & P-Value \\
\hline \\[-1.8ex] 
& \multicolumn{5}{l}{Panel B: Covariate Balance on Individual Characteristics} \\
\hline \\[-1.8ex]
Altruist &  0.63 &  0.68 &  0.62 & 0.239 & 0.788 \\ 
Age & 25.01 & 26.00 & 26.90 & 0.754 & 0.472 \\ 
Religiosity &  0.53 &  0.38 &  0.45 & 1.487 & 0.228 \\ 
Pol. Engagement &  0.69 &  0.65 &  0.68 & 0.163 & 0.849 \\ 
Soc. Engagement &  0.59 &  0.62 &  0.67 & 0.454 & 0.636 \\ 
Above 2 Min. Wage &  0.34 &  0.35 &  0.29 & 0.336 & 0.715 \\ 
FB Popularity &  0.53 &  0.56 &  0.58 & 0.175 & 0.840 \\ 
\hline \\[-1.8ex] 
& \multicolumn{5}{l}{Panel C: Unmatched Covariate Balance on Territorial Assignment} \\
\hline \\[-1.8ex] 
Number of Houses & 276.41 & 243.58 & 255.20 &  2.368 & 0.098 \\ 
Number of Households & 885.67 & 750.34 & 785.18 &  3.765 & 0.026 \\ 
Avg. Household Size &   3.21 &   3.07 &   3.07 &  6.566 & 0.002 \\ 
Log Avg. Income &   6.47 &   6.86 &   6.85 &  9.213 & 0.000 \\ 
Cases Before Treatment &  13.78 &   5.94 &   4.66 & 15.638 & 0.000 \\ 
\hline \\[-1.8ex] 
& \multicolumn{5}{l}{Panel D: Matched Covariate Balance on Territorial Assignment} \\
\hline \\[-1.8ex] 
Number of Houses & 276.41 & 282.04 & 281.19 & 0.068 & 0.935 \\ 
Number of Households & 885.67 & 903.04 & 900.39 & 0.060 & 0.942 \\ 
Avg. Household Size &   3.21 &   3.21 &   3.20 & 0.008 & 0.992 \\ 
Log Avg. Income &   6.47 &   6.45 &   6.45 & 0.084 & 0.920 \\ 
Cases Before Treatment &  13.78 &  10.24 &  10.77 & 1.443 & 0.240 \\ 
\hline
Pairs & 34 & 34 & 35 &  &  \\ 
N & 68 & 68 & 70 &  &  \\
\hline\hline \\[-1.8ex] 
\end{tabular}
\end{table}

\normalsize

The intervention took place on the 5^th^ of May, 2018. We randomly assigned
subjects into three groups and directed each group to separate headquarters in
order to avoid spillovers. In the control group ($n = 68$), we told participants
that they would receive a flat payment regardless of their performance (BRL 110
/ USD 20 per subject). In the individual treatment group ($n = 68$), we informed
subjects that we would rank individual performances and double the compensation
for those who performed above the sample median (BRL 220 / USD 50 per subject).
In the collective treatment group ($n = 70$), we explained that we would rank
the pair performance and double the payment for teams who presented results
better than the sample median (BRL 220 / USD 50 per subject). Participants
worked in pairs and were aware of other groups, but they did not know about the
different treatment conditions. The teams received a leaflet with a map of the
area they should cover. The route comprised of two to three blocks and contained
around 120 houses.

A typical visit to a household consisted of the following steps. First,
participants rang the doorbell and explained that the municipality had
experienced a dengue fever outbreak. They instructed dwellers on how to lower
dengue incidence and handed them a leaflet with information about dengue
prevention practices. Then, subjects asked whether they were allowed to inspect
house yard. If granted permission, they entered the household and searched for
clean breeding sites, such as pots filled with clear water, and recipients
containing larvae. Upon exterminating a breeding site (pouring out the
accumulated water, for example), subjects had to report their task by taking a
photo on their cellphones. When they discovered larvae, they had to record a
video showing the larvae before exterminating it. At the end of their visit,
they had to take a picture of the household to account for their presence in the
location.

This study evaluates two outcomes. The first outcome is field productivity of
the health care workers. We measure worker productivity by coding cellphone data
collected during the intervention into four indicators: 1) the number of houses
visited; 2) the number of houses visited within less than two minutes; 3) the
number of breeding sites removed and cleaned; 4) the number of exterminated
larvae. We analyse the data using ordinary least squares with robust standard
errors [@imbens2016robust]. We estimate the statistical power of the study
assuming a treatment effect of 0.15 standard deviations.

The second outcome of interest is dengue incidence in the municipality of Rio
Verde measured 4, 8, 12, and 16 days after the intervention. The Rio Verde
Mayor's Office provided us with its annual health care report, which contains
the dates and personal information of patients who had dengue fever and received
hospital treatment from January to September 2018. We georeferenced the home
addresses of the patients and compared the incidence map with the distribution
of our intervention. We also compare dengue incidence in each of the treatment
statuses with incidence figures in the control group. The unit of analysis is
the census tract in both models. We employ a differences-in-differences
estimator to assess the treatment effects [@angrist2008mostly].

We pre-registered the experiment on the Evidence in Governance and Policy (EGAP)
website (<https://osf.io/6q8vu/>) and received IRB approval from New York
University (IRB-FY2017-17) and Fundação Getulio Vargas (IRB-01/2017). The
funders of this study had no influence over the formulation, organisation,
analysis, and interpretation of the outcomes. Health care workers gave written
consent upon filling the Qualtrics questionnaire. During the intervention, the
occupants of households inspected for breeding sites and larvae gave oral
consent. For more information on the experimental design, sample
characteristics, exclusion criteria, and estimation methods, please refer to the
Supplementary Material.

# Results
\label{sec:results}

We find that financial incentives significantly increase the field productivity
of health care workers fighting *Aedes aegypti*. Subjects in both the
individual and the collective bonus groups locate more mosquito breeding sites
and exterminate more larvae than participants in the control condition. Our
results also show the intervention decreases city-wide levels of dengue
incidence, but the coefficients only reach statistical significance when we
combine the two treatments together and contrast them with the control group.
Overall, the findings suggest that monetary payments are an effective method of
improving health worker performance, yet more research is needed to assess
whether higher labour productivity leads to fewer dengue-related
hospitalisations.

## Field Productivity
\label{sub:productivity}

We have four indicators to measure health worker productivity. The first is the
number of houses visited during the intervention. The second indicator is the
number of houses in which the visits lasted for less than two minutes. The
third dependent variable is the number of *A. aegypti* breeding sites removed
and cleaned. Our fourth variable indicates the number of mosquito larvae
exterminated. Table \ref{tab:productivity} summarises the experimental results.

\begin{table}[ht] 
  \centering 
  \small
  \caption{Field Productivity} 
  \label{tab:productivity} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent Variable:}} \\ 
\cline{2-5} 
\\[-1.8ex] & Houses & Houses & Breeding & Larvae \\ 
& Visited & Visited ($<$ 2min.) & Sites & Exterminated \\ 
\hline \\[-1.8ex] 
 Individual Bonus & $-$9.879$^{**}$ & $-$8.044$^{**}$ & 25.118$^{***}$ & 0.022 \\ 
  & (4.906) & (4.018) & (5.251) & (0.058) \\ 
  & & & & \\ 
 Collective Bonus & $-$7.554$^{*}$ & $-$6.479$^{*}$ & 21.512$^{***}$ & 0.163$^{**}$ \\ 
  & (4.428) & (3.571) & (4.049) & (0.076) \\ 
  & & & & \\ 
\hline \\[-1.8ex] 
Observations & 197 & 197 & 197 & 197 \\ 
Residual SE & 27.301 & 22.153 & 30.545 & 0.425 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{r}{Robust standard errors in parentheses. $^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

Subjects who received monetary incentives, either alone or in pairs, visited
fewer houses than those who did not earn financial rewards. While the result
may seem counter-intuitive, it actually attests the effectiveness of the
treatment. This is because we offered incentives for subjects to exterminate
larvae and find breeding sites, but not to visit a greater number of
households. As a result, participants in the two treatment groups did a
thorough job in each household, which explains the lower number of visits
during the allotted time. We find that subjects in the individual and in the
collective bonus groups visited $-9.879$ (CI: [XXX, XXX], $p-$value $= 0.XXX$)
and $-7.554$ (CI: [XXX, XXX], $p-$value $= 0.XXX$) households, respectively.

Our second indicator provides further evidence in favour of our argument. We
see that the two groups which earned performance bonuses spent more time in the
households, as their number of houses visited for less than two minutes is
significantly lower than those of the control condition. On average,
participants in the individual bonus treatment arm visited $-8.044$ (CI: [XXX,
XXX], $p-$value $= 0.XXX$) fewer houses than the control, and those in the
collective bonus group visited $-6.479$ (CI: [XXX, XXX], $p-$value $= 0.XXX$).

Incentivised teams find and clean more breeding sites than non-incentivised
ones. Individually incentivised health workers find 25.118 (CI: [XXX, XXX],
$p-$value $= 0.XXX$) more breeding sites than the non-incentivised control
group. The effect for those in the collective bonus group is 21.512 (CI: [XXX,
XXX], $p-$value $= 0.XXX$), slightly smaller than for the first treatment.

Collective incentivized teams are more effective in finding larvae. In our
study, collectively incentivized subjects have 16.3\% more chance to find
larvae than non incentivized workers. Finding larvae is harder, and the
detection chances profit from teamwork. If teams are cooperating, searching the
house becomes more natural, and they can coordinate to do a more thorough job.
Additionally, the detection tends to be excludable: one individual finding
larvae makes the chances of keep finding larvae in the household considerably
smaller. Collective bonus is not excludable within the teams, making them
cooperate more to get the reward.

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{5pt}