### 2021-02-25 manuscript/article.Rmd
add ethics statement, small edits

---
title:  |
        | Financial Incentives and Healthcare Provision: Evidence from an Experimental *Aedes aegypti* Control Programme in Brazil
author:
  - Danilo Freire[^freire]
  - Umberto Mignozzetti[^mignozzetti]
date: \today
thanks: "We thank Taylor Boas, Saad Gulzar, Cathy Hafer, Daniel Hidalgo, Dimitri Landa, Rebecca Morton, Adam Przeworski, Catarina Roman, Renard Sexton, David Skarbek, Denis Stukal, and participants at EGAP, MPSA, NYU PE Seminar, and SPSA for their valuable comments. We also thank the FGV Applied Research Centre for funding this research. Each named author has equally contributed to conducting the underlying research and drafting this manuscript. To the best of our knowledge, the authors have no conflict of interest, financial or otherwise. Replication materials are available at <http://github.com/danilofreire/incentives-healthcare>."
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: unsrtnat
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: template.latex
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# If you need to install any package while knitting the document
r <- getOption("repos")
r["CRAN"] <- "https://cran.rstudio.com/"
options(repos = r)

# Install required packages
pkgs <- c("tidyverse", "magrittr", "lmtest",
          "estimatr", "sandwich", "stargazer",
          "gridExtra", "WeightIt", "robustHD",
          "kableExtra", "lfe", "fixest")

installIfNot <- function(x) {
  if(x %in% rownames(installed.packages()) == FALSE)
    install.packages(x, dependencies = T)
}
invisible(lapply(pkgs, installIfNot))

# Load packages
invisible(lapply(pkgs, require, character.only = T))
rm(pkgs, installIfNot, r)

# Load data
load("../data/incentives_healthcare.RData")
```

[^freire]: Independent Researcher,
[`danilofreire@gmail.com`](mailto:danilofreire@gmail.com),
<http://danilofreire.github.io>.

[^mignozzetti]: Visiting Assistant Professor, Quantitative Theory and Methods
Department, Emory University, 532 North Kilgo Circle, 4^th^ Floor, Atlanta, GA
30322, USA,
[`umberto.mignozzetti@emory.edu`](mailto:umberto.mignozzetti@emory.edu),
<http://umbertomig.com>. Corresponding author.

\newpage

\begin{abstract}

\footnotesize

\noindent \textbf{Background}. Mosquito control is the most effective means of
reducing \textit{Aedes aegypti} infections worldwide. In many developing
countries, however, vector management programmes fail to reach their goals due
to low worker productivity. Research suggests that financial incentives may
increase the productivity of health personnel, yet there is little evidence
about the impact of monetary rewards on \textit{A. aegypti}-reduction
strategies. We evaluated whether individual and collective financial incentives
improve the performance of healthcare workers fighting \textit{A. aegypti}, as
well as their effect on city-level numbers of dengue hospitalisations.

\noindent \textbf{Methodology/Principal findings}. We hired and trained
subjects to visit households, find \textit{A. aegypti} breeding sites, and
eliminate mosquito larvae in the city of Rio Verde, Brazil. We randomly
assigned workers into three groups. The control group received a flat
compensation for their tasks, while workers in the two treatment groups received
individual and collective monetary bonuses, respectively. Financial rewards
increased the number of cleaned breeding sites in both treatment groups
(individual and team bonuses), and the collective treatment also improved larvae
extermination. The intervention lowered dengue hospitalisations in 10.3\%, but
the result was not consistent across all model specifications.

\noindent \textbf{Conclusions/Significance}. \textit{A. aegypti} control
programmes may benefit from alternative compensation schemes, especially when
provided to teams. For this strategy to succeed, financial incentives have to
be distributed widely as their aggregate effect is limited. More research is
needed to assess whether higher worker productivity decreases dengue
hospitalisations.

\end{abstract}

\renewcommand{\abstractname}{Author Summary}
\begin{abstract}

\footnotesize

\noindent Diseases transmitted by the \textit{Aedes aegypti} mosquito, such as
chikungunya, dengue, yellow fever, and Zika, continue to affect thousands of
people per year. As there are no safe vaccines for most of these infections,
insecticide spraying and breeding site elimination are the best means to fight
the mosquito. In several developing countries, which host the majority of
\textit{A. aegypti} infections, anti-mosquito campaigns are carried out
inconsistently, thus it is crucial to find ways to improve the productivity of
healthcare workers in charge of these tasks. We designed a randomised field
experiment that provided individual and collective financial incentives to
healthcare agents in a Brazilian city, and we tested the effect of monetary
rewards on their productivity and on city-level dengue hospitalisations. We
find that financial bonuses improved the number of cleaned breeding sites in
both treatment groups (individual and team incentives) and that the collective
treatment also improved larvae extermination. The impact of our treatment on
city-level hospitalisations was not consistent across all specifications. In
sum, financial incentives may be used to boost field productivity in
anti-\textit{A. aegypti} programmes, but further research is required to
evaluate how healthcare worker productivity impacts dengue outcomes.

\vspace{.3cm} 
\noindent \textbf{Keywords}: Aedes aegypti, dengue, financial
incentives, mosquito control, preventive healthcare, randomised field
experiment, worker productivity 
\end{abstract}

\normalsize

\newpage

# Introduction
\label{sec:intro}

The *Aedes aegypti* mosquito is the major worldwide vector of several arboviral
diseases, including chikungunya, dengue, yellow fever, and Zika
[@boas2019electoral; @burt2012chikungunya; @jansen2010dengue; @leta2018global;
@thangamani2016vertical]. Over the last decades, arboviral incidence has
increased exponentially, with estimates suggesting that dengue is endemic in
over 100 countries and causes 400 million infections and 22 thousand deaths
each year [@bhatt2013global; @cdc2020dengue]. Since there are no safe vaccines
for most *Aedes*-transmitted diseases [@world2017dengue], we have seen a number
of dengue and Zika outbreaks in Asia, the Caribbean, South America, and the
United States [@carlson2016ecological; @leta2018global]. Chikungunya virus,
previously endemic only to Africa, has already been reported in every region of
the world [@grandadam2011chikungunya; @kraemer2015global; @mayer2017emergence].

Mosquito control is the primary strategy to prevent *Aedes*-borne infections
[@beatty2010best; @eisen2009proactive]. The World Health Organization
recommends the cleaning of containers and the use of insecticide to reduce
vector breeding, especially in settings where human-vector contacts are
frequent, such as schools, hospitals, and workplaces [@WHO2012global;
@world2020denguecontrol]. However, breeding site elimination and insecticide
spraying are labour-intensive processes, which are often carried out
inconsistently by developing states [@paz2015dengue; @siddiqui2016use;
@WHO2012global]. In this regard, improving the performance of their healthcare
agencies is essential to the effectiveness of anti-*Aedes* programmes.

Research shows that financial incentives may increase the productivity of
healthcare workers [@clemens2014physicians; @muthmainnah2017improving;
@ulrich2005does]. Nevertheless, systematic reviews indicate that the internal
validity of the available evidence is limited due to methodological
shortcomings, as several studies suffer from selection bias and inadequate
experimental blinding [@bassani2013financial; @flodgren2011overview;
@scott2011effect]. The reviews also point out that it remains unclear which
rewarding schemes have a stronger impact on healthcare worker performance
[@fainman2020design] and whether monetary incentives have any sustained effects
on patient outcomes [@jha2012long; @langdown2014use].

In this paper, we explore how financial incentives affect the productivity of
healthcare personnel involved in *A. aegypti* control and how they impact
dengue hospitalisations. We ran a randomised field experiment in the city of
Rio Verde, Brazil, which has experienced a peak in dengue infections in recent
years. We hired and trained teams of healthcare workers to perform simplified
procedures designed to exterminate *A. aegypti* breeding sites. Our experiment
consisted of two treatment arms plus a control condition. We offered individual
or collective bonuses to those allocated in the two treatment groups, while the
control group received a fixed compensation for completing their tasks. We then
measured how the different reward schemes affected the number of breeding sites
discovered, larvae exterminated, and city-wide dengue incidence.

We find that monetary rewards increase the number of cleaned breeding sites in
both treatment conditions (individual and team bonuses), and that collective
financial incentives also improve larvae extermination. Our results show that
the control group has a high number of houses visited in less than two minutes
apart from each other, which suggests that workers may cheat in the absence of
monetary incentives. When we combine both treatments, we also find that the
intervention decreases dengue incidence in 10.3%, but the results are not
robust. In summary, we provide experimental evidence that performance bonuses
enhance public service delivery and that small financial incentives may improve
patient outcomes and reduce hospitalisations in *Aedes*-endemic countries.

# Methods
\label{sec:methods}

We conducted a randomised field trial in Rio Verde, a municipality in the
Brazilian Midwestern state of Goiás. In 2017, Rio Verde had 229,000 inhabitants
and 1,156 confirmed dengue cases, one of highest counts in the state
[@goias2020]. We used Facebook advertisements to recruit potential
participants. The advertisements directed applicants to Google Forms, where we
asked for their contact information. We sent a pre-treatment demographic survey
to every applicant who provided a valid email address. At this stage, we
excluded all individuals under 18 years old and invited the remaining
applicants to our training session, which we ran on the 4^th^ of May, 2018. The
meeting involved a lecture on *A. aegypti* vector biology and practical
instructions on how to find and exterminate mosquito breeding sites. Those who
did not attend the session were also excluded from the study.

Table \ref{tab:covbal} displays the attrition rates and the covariance balance
tests for our experiment. The last two tables report the values for our
territorial assignment before and after applying propensity score matching
[@rosenbaum1985constructing] on census tract characteristics. We observe no
significant differences regarding attrition rates, indicating that the
treatment was correctly randomised. Participant demographics are also similar
across all groups, which suggests that the matching technique successfully
balanced the territorial assignment.

\begin{table}[ht] \centering
\small
  \caption{Covariate Balance on Individual Characteristics}
  \label{tab:covbal}
\begin{tabular}{@{\extracolsep{5pt}} lcccccc}
\\[-1.8ex]\hline
\hline \\[-1.8ex]
Variable & Attrition & Control & Individual & Collective & F-Statistic & P-Value \\
\hline \\[-1.8ex]
& & \multicolumn{5}{l}{Panel A: Attrition Pattern Test} \\
\hline \\[-1.8ex]
```{r, echo = FALSE, results='asis', cache=T}
tabcov <- data.frame()
pretreat$assign2 <- ifelse(is.na(pretreat$assignment)|
                             pretreat$participating!='Participating',
                           'Attrition', pretreat$assignment)
pretreat$assign2 <- factor(pretreat$assign2, levels = c('Attrition',
                                                        'No bonus',
                                                        'Individual',
                                                        'Collective'))
x = tapply(pretreat$Age,
           pretreat$assign2,
           mean)
y = summary(aov(Age~assign2,
                data = pretreat))
tab <- data.frame(Var = 'Age',
                  `Atrition` = round(x[1],2),
                  `NoBonus` = round(x[2],2),
                  `Individual` = round(x[3],2),
                  `Collective` = round(x[4],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)
x = tapply(pretreat$Female,
           pretreat$assign2,
           mean)
y = summary(aov(Female~assign2,
                data = pretreat))
tab <- data.frame(Var = 'Female',
                  `Atrition` = round(x[1],2),
                  `NoBonus` = round(x[2],2),
                  `Individual` = round(x[3],2),
                  `Collective` = round(x[4],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)
x = tapply(pretreat$HasVehicle,
           pretreat$assign2,
           mean)
y = summary(aov(HasVehicle~assign2, data = pretreat))
tab <- data.frame(Var = 'Has Vehicle',
                  `Atrition` = round(x[1],2),
                  `NoBonus` = round(x[2],2),
                  `Individual` = round(x[3],2),
                  `Collective` = round(x[4],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)
x = round(tapply(rep(1, length(pretreat$assign2)),
                 pretreat$assign2,
                 sum), digits = 0)
tab <- data.frame(Var = 'N',
                  `Atrition` = round(x[1],2),
                  `NoBonus` = round(x[2],2),
                  `Individual` = round(x[3],2),
                  `Collective` = round(x[4],2),
                  FVal = NA,
                  Pvalue = NA)
tabcov <- rbind(tabcov, tab)
row.names(tabcov) <- NULL
for (i in 1:dim(tabcov)[1]) {
  s <- paste(tabcov[i,], collapse = ' & ')
  s <- gsub('NA', '', paste(tabcov[i,], collapse = ' & '))
  cat(paste(s, ' \\\\ ', sep = ''), '\n')
}
```
\hline \\[-1.8ex]
\end{tabular}
\begin{tabular}{@{\extracolsep{5pt}} lccccc}
Variable & Control & Individual & Collective & F-Statistic & P-Value \\
\hline \\[-1.8ex]
& \multicolumn{5}{l}{Panel B: Covariate Balance on Individual Characteristics} \\
\hline \\[-1.8ex]
```{r, echo = FALSE, results='asis', cache=T}
tabcov <- data.frame()
pretreat$assign3 <- ifelse(is.na(pretreat$assignment)|
                             pretreat$participating!='Participating',
                           NA, pretreat$assignment)
pretreat$assign3 <- factor(pretreat$assign3, levels = c('No bonus',
                                                        'Individual',
                                                        'Collective'))
aux <- filter(pretreat, !is.na(assign3))
# Age
x = tapply(aux$Age,
           aux$assign3,
           mean)
y = summary(aov(Age~assign3,
                data = aux))
tab <- data.frame(Var = 'Age',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# Altruism
x = tapply(aux$aboveMedianAltr,
           aux$assign3,
           mean)
y = summary(aov(aboveMedianAltr~assign3,
                data = aux))
tab <- data.frame(Var = 'Altruist',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# Religiosity
x = tapply(aux$relig,
           aux$assign3,
           mean)
y = summary(aov(relig~assign3,
                data = aux))
tab <- data.frame(Var = 'Religiosity',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# Political Engagement
x = tapply(aux$polengage,
           aux$assign3,
           mean)
y = summary(aov(polengage~assign3,
                data = aux))
tab <- data.frame(Var = 'Pol. Engagement',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# Social Engagement
x = tapply(aux$socengage,
           aux$assign3,
           mean)
y = summary(aov(socengage~assign3,
                data = aux))
tab <- data.frame(Var = 'Soc. Engagement',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# Above Two Minimum Wage Family
x = tapply(aux$aboveTwoMinWageFam,
           aux$assign3,
           mean)
y = summary(aov(aboveTwoMinWageFam~assign3,
                data = aux))
tab <- data.frame(Var = 'Above Two Min. Wage',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# Above Median Facebook popularity
x = tapply(aux$FBpop,
           aux$assign3,
           mean)
y = summary(aov(FBpop~assign3,
                data = aux))
tab <- data.frame(Var = 'Facebook Popularity',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# N
x = round(tapply(rep(1, length(aux$assign3)),
                 aux$assign3,
                 sum), digits = 0)
row.names(tabcov) <- NULL
for (i in 1:dim(tabcov)[1]) {
  s <- paste(tabcov[i,], collapse = ' & ')
  s <- gsub('NA', '', paste(tabcov[i,], collapse = ' & '))
  cat(paste(s, ' \\\\ ', sep = ''), '\n')
}
```
\hline \\[-1.8ex]
& \multicolumn{5}{l}{Panel C: Unmatched Covariate Balance on Territorial Assignment} \\
\hline \\[-1.8ex]
```{r, echo = FALSE, results='asis', cache=T}
## Census Sectors characteristics
tabcov <- data.frame()

# Number of Houses
x = tapply(dfr2$NHouses,
           dfr2$D1,
           mean)
y = summary(aov(NHouses~D1, data = dfr2))
tab <- data.frame(Var = 'Number of Houses',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# Age
x = tapply(dfr2$NHouseholds,
           dfr2$D1,
           mean)
y = summary(aov(NHouseholds~D1, data = dfr2))
tab <- data.frame(Var = 'Number of Households',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# AvgHouseholdSize
x = tapply(dfr2$AvgHouseholdSize,
           dfr2$D1,
           mean)
y = summary(aov(AvgHouseholdSize~D1, data = dfr2))
tab <- data.frame(Var = 'Avg. Household Size',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# Avg. Income Households
x = tapply(dfr2$logAvgIncomeHousehold,
           dfr2$D1,
           mean)
y = summary(aov(logAvgIncomeHousehold~D1, data = dfr2))
tab <- data.frame(Var = 'Log Avg. Income',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# Cases Before
x = tapply(dfr2$plac,
           dfr2$D1,
           mean)
y = summary(aov(plac~D1, data = dfr2))
tab <- data.frame(Var = 'Cases Before Treatment',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)
row.names(tabcov) <- NULL
## re-weighting
dfr2$weightsD1 = weightit(D1~NHouses+
                    NHouseholds+AvgHouseholdSize+
                    logAvgIncomeHousehold+plac,
                  data = dfr2, focal = "No bonus",
                  method = 'cbps',
                  estimand = "ATT")$weights

dfr$weightsD1 <- weightit(D1~NHouses+
                    NHouseholds+AvgHouseholdSize+
                    logAvgIncomeHousehold+plac,
                  data = dfr, focal = "Control",
                  method = 'cbps',
                  estimand = "ATT")$weights

dfr$weightsD2 <- weightit(D2~NHouses+
                            NHouseholds+AvgHouseholdSize+
                            logAvgIncomeHousehold+plac,
                          data = dfr, focal = "No intervention",
                          method = 'cbps',
                          estimand = "ATT")$weights
for (i in 1:dim(tabcov)[1]) {
  s <- paste(tabcov[i,], collapse = ' & ')
  s <- gsub('NA', '', paste(tabcov[i,], collapse = ' & '))
  cat(paste(s, ' \\\\ ', sep = ''), '\n')
}
```
\hline \\[-1.8ex]
& \multicolumn{5}{l}{Panel D: Matched Covariate Balance on Territorial Assignment} \\
\hline \\[-1.8ex]
```{r, echo = FALSE, results='asis', cache=T}
tabcov <- data.frame()
# Number of Houses
x = dfr2 %>%
  group_by(D1) %>%
  summarise(x = weighted.mean(NHouses, weightsD1),
            .groups = 'drop')
x = x$x
y = summary(aov(NHouses~D1, data = dfr2, weights=dfr2$weightsD1))
tab <- data.frame(Var = 'Number of Houses',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# Number of households
x = dfr2 %>%
  group_by(D1) %>%
  summarise(x = weighted.mean(NHouseholds, weightsD1),
            .groups = 'drop')
x = x$x
y = summary(aov(NHouseholds~D1, data = dfr2, weights=dfr2$weightsD1))
tab <- data.frame(Var = 'Number of Households',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# AvgHouseholdSize
x = dfr2 %>%
  group_by(D1) %>%
  summarise(x = weighted.mean(AvgHouseholdSize, weightsD1),
            .groups = 'drop')
x = x$x
y = summary(aov(AvgHouseholdSize~D1, data = dfr2,
                weights=dfr2$weightsD1))
tab <- data.frame(Var = 'Avg. Household Size',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# Avg. Income Households
x = dfr2 %>%
  group_by(D1) %>%
  summarise(x = weighted.mean(logAvgIncomeHousehold, weightsD1),
            .groups = 'drop')
x = x$x
y = summary(aov(logAvgIncomeHousehold~D1, data = dfr2,
                weights=dfr2$weightsD1))
tab <- data.frame(Var = 'Log Avg. Income',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)

# Cases Before
x = dfr2 %>%
  group_by(D1) %>%
  summarise(x = weighted.mean(plac, weightsD1),
            .groups = 'drop')
x = x$x
y = summary(aov(plac~D1, data = dfr2, weights=dfr2$weightsD1))
tab <- data.frame(Var = 'Cases Before Treatment',
                  NoBonus = round(x[1],2),
                  Individual = round(x[2],2),
                  Collective = round(x[3],2),
                  FVal = round(y[[1]][1,4], 3),
                  Pvalue = round(y[[1]][1,5], 3))
tabcov <- rbind(tabcov, tab)
row.names(tabcov) <- NULL
for (i in 1:dim(tabcov)[1]) {
  s <- paste(tabcov[i,], collapse = ' & ')
  s <- gsub('NA', '', paste(tabcov[i,], collapse = ' & '))
  cat(paste(s, ' \\\\ ', sep = ''), '\n')
}
```
\hline
Pairs & 29 & 32 & 37 &  &  \\
N & 58 & 64 & 74 &  &  \\
\hline\hline \\[-1.8ex]
\end{tabular}
\end{table}

\normalsize

The intervention took place on the 5^th^ of May, 2018. We randomly assigned
subjects into three groups and directed each group to separate headquarters to
avoid spillovers. In the control group (n = `r as.numeric(table(prodfr$D)[1])`),
we told participants that they would receive a flat payment regardless of their
performance (BRL 110 / USD 25 per subject). In the individual treatment group
(n = `r as.numeric(table(prodfr$D)[2])`), we informed participants that we would
rank each one of them separately and double the compensation for those who
performed above the sample median (BRL 220 / USD 50 per subject). In the
collective treatment group (n = `r as.numeric(table(prodfr$D)[3])`), we
explained that we would rank the performance of each group and double the
payment for those teams whose results were higher than the sample median (BRL
220 / USD 50 per subject). Participants worked in pairs and were aware of other
groups, but they did not know about the different treatment conditions. The
teams received a leaflet with a map of the area they should cover. Each route
included around 120 houses in two or three blocks.

A typical household visit consisted of the following steps. First, participants
rang the doorbell and explained that the municipality had experienced a dengue
fever outbreak. They instructed dwellers on how to lower dengue incidence and
handed them a leaflet with information about dengue prevention practices. Then,
subjects asked whether they were allowed to inspect the resident's house yard.
If granted permission, they entered the household and searched for clean
breeding sites, such as pots filled with clear water, and for recipients
containing larvae. Upon exterminating a breeding site (pouring out the
accumulated water, for example), subjects had to report their task by taking a
photo on their cellphones. When they discovered larvae, participants had to
record a short video showing the larvae before exterminating them. At the end
of each visit, subjects had to take a picture of the household to account for
their presence in the location.

This study evaluates two outcomes. The first outcome is the field productivity
of healthcare providers. We measure worker productivity by coding cellphone
data collected during the intervention into four indicators: 1) the number of
breeding sites removed and cleaned; 2) the number of exterminated larvae; 3)
the number of houses visited; 4) the number of houses visited within less than
two minutes. We analyse the data using ordinary least squares with robust
standard errors [@imbens2016robust]. We estimate the statistical power of the
study assuming a treatment effect of 0.15 standard deviations.

The second outcome of interest is dengue incidence in the municipality of Rio
Verde measured 4, 8, 12, and 16 days after the intervention. The Rio Verde
Mayor's Office provided us with its annual healthcare report, which contains
the dates and personal information of patients who had dengue fever and
received hospital treatment from January to September 2018. We georeferenced
the home addresses of the patients and compared the incidence map with the
distribution of our intervention. We also compare dengue incidence in each of
the treatment statuses with incidence figures in the control group. In both
models, the unit of analysis is the census tract. We employ a
differences-in-differences estimator to measure the treatment effects
[@angrist2008mostly].

## Ethics Statement

We pre-registered the experiment on the Evidence in Governance and Policy
(EGAP) website (#20180504AA, <https://osf.io/6q8vu/>), and received IRB
approval from New York University (IRB-FY2017-17) and Fundação Getulio Vargas
(IRB-01/2017). The funders of this study had no influence over the formulation,
organisation, analysis, and interpretation of the outcomes. Healthcare workers
gave written consent upon filling the pre-treatment questionnaire. During the
intervention, the occupants of households inspected for breeding sites and
larvae gave verbal consent. For more information about the experimental design, please refer to the Supplementary Material.

# Results
\label{sec:results}

We find that financial incentives significantly increase the field productivity
of healthcare workers fighting *Aedes aegypti*. Subjects in both the
individual and the collective bonus groups locate more mosquito breeding sites
and exterminate more larvae than participants in the control condition. Our
results also show the intervention decreases city-wide levels of dengue
incidence, but the coefficients only reach statistical significance when we
combine the two treatments and contrast them with the control group. Overall,
the findings suggest that monetary payments are an effective method of
improving health worker performance, yet more research is needed to assess
whether higher labour productivity leads to fewer dengue-related
hospitalisations.

## Field Productivity
\label{sub:productivity}

```{r, echo = FALSE, results='asis', warning=FALSE, cache=T}
## Productivity outcome
library(lmtest)
mprod <- list()
mprod[[1]] <- lm(nHouse ~ D, data = prodfr)
mprod[[2]] <- lm(nHouseLessTwoMin ~ D, data = prodfr)
mprod[[3]] <- lm(nBreed ~ D, data = prodfr)
mprod[[4]] <- lm(nLarvae ~ D, data = prodfr)
```

Table \ref{tab:productivity} summarises the experimental results. 
Subjects who receive monetary incentives find and clean more breeding sites
than those who earned a flat compensation fee. Workers in the individual
treatment group find `r round(as.numeric(coef(mprod[[3]])[2]), 3)` (CI: [`r round(coefci(mprod[[3]], vcov = vcovHC(mprod[[3]], type = "HC1"), level = 0.9), digits=3)[2,1]`, `r round(coefci(mprod[[3]], vcov = vcovHC(mprod[[3]], type = "HC1"), level = 0.9), digits=3)[2,2]`]) more breeding sites than the
non-incentivised control group. The effect for those in the collective bonus
group is `r round(as.numeric(coef(mprod[[3]])[3]), 3)` (CI: [`r round(coefci(mprod[[3]], vcov = vcovHC(mprod[[3]], type = "HC1"), level = 0.9), digits=3)[3,1]`, `r round(coefci(mprod[[3]], vcov = vcovHC(mprod[[3]], type = "HC1"), level = 0.9), digits=3)[3,2]`]), slightly smaller than for the individual treatment. This finding suggests that the treatment does improve the performance of healthcare workers.

```{r, echo = FALSE, results='asis', warning=FALSE, cache=T}
## Productivity outcome
mprod <- list()
mprod[[1]] <- lm(nHouse ~ D, data = prodfr)
mprod[[2]] <- lm(nHouseLessTwoMin ~ D, data = prodfr)
mprod[[3]] <- lm(nBreed ~ D, data = prodfr)
mprod[[4]] <- lm(nLarvae ~ D, data = prodfr)
stargazer(mprod[[3]], mprod[[4]],
          mprod[[1]], mprod[[2]],
          se = starprep(mprod[[3]], mprod[[4]],
                        mprod[[1]], mprod[[2]],
                        se_type = "HC1"),
          dep.var.labels = c('\\shortstack{Breeding Sites \\\\ Exterminated}',
                             '\\shortstack{Larvae \\\\ Exterminated}',
                             '\\shortstack{Houses \\\\ Visited}',
                             '\\shortstack{Houses Visited \\\\ in Less than \\\\ Two Minutes}'),
          keep.stat = c('ser', 'n'),
          notes = c('Robust standard errors in parenthesis.'),
          keep = c('DIndividual', 'DCollective'),
          covariate.labels = c('Individual Bonus',
                               'Collective Bonus'),
          title = 'Field Productivity', header = FALSE,
          label = 'tab:productivity')
```

\newpage

Teams who received collective bonuses are the most effective in finding
larvae. Subjects who worked under this treatment status have a `r round(as.numeric(coef(mprod[[4]])[3]*100),1)`% (CI: [`r round(coefci(mprod[[4]], vcov = vcovHC(mprod[[4]], type = "HC1"), level = 0.9), digits=3)[3,1]*100`%, `r round(coefci(mprod[[4]], vcov = vcovHC(mprod[[4]], type = "HC1"), level = 0.9), digits=3)[3,2]*100`%]) higher
likelihood of finding larvae than those in the control group. Locating mosquito
larvae is a difficult task, thus the probability of successfully completing the
task increases when subjects work in teams. Furthermore, our collective bonus
scheme was not excludable within the teams, thus healthcare workers had an
incentive to cooperate and receive higher rewards.

Subjects who received monetary incentives visited fewer houses than those who
did not earn financial rewards. The result is statistically significant for the
individual treatment group. While the finding seems counter-intuitive, it
actually attests the effectiveness of the treatment. This is because we offered
incentives for subjects to exterminate larvae and find breeding sites, not to
visit a greater number of households. As a result, participants in the 
treatment groups did a thorough job in each household, which explains the lower
number of visits during the allotted time. We find that subjects in the
individual and in the collective bonus groups visited `r round(as.numeric(coef(mprod[[1]])[2]), 3)` (CI: [`r round(coefci(mprod[[1]], vcov = vcovHC(mprod[[1]], type = "HC1"), level = 0.9), digits=3)[2,1]`, `r round(coefci(mprod[[1]], vcov = vcovHC(mprod[[1]], type = "HC1"), level = 0.9), digits=3)[2,2]`]) and `r round(as.numeric(coef(mprod[[1]])[3]), 3)` (CI: [`r round(coefci(mprod[[1]], vcov = vcovHC(mprod[[1]], type = "HC1"), level = 0.9), digits=3)[3,1]`, `r round(coefci(mprod[[1]], vcov = vcovHC(mprod[[1]], type = "HC1"), level = 0.9), digits=3)[3,2]`]) households, respectively.

Our last indicator, the number of houses visited within less than two minutes, provides further evidence in favour of our argument. On average,
participants in the individual bonus treatment arm visited `r round(as.numeric(coef(mprod[[2]])[2]), 3)` (CI: [`r round(coefci(mprod[[2]], vcov = vcovHC(mprod[[2]], type = "HC1"), level = 0.9), digits=3)[2,1]`,
`r round(coefci(mprod[[2]], vcov = vcovHC(mprod[[2]], type = "HC1"), level = 0.9), digits=3)[2,2]`]) houses than the control, and those in the collective bonus group visited
`r round(as.numeric(coef(mprod[[2]])[3]), 3)` (CI: [`r round(coefci(mprod[[2]], vcov = vcovHC(mprod[[2]], type = "HC1"), level = 0.9), digits=3)[3,1]`, `r round(coefci(mprod[[2]], vcov = vcovHC(mprod[[2]], type = "HC1"), level = 0.9), digits=3)[3,2]`]). Again, the result for the individual bonus group is statistically significant at the 10% level.

## Disease Incidence
\label{sub:incidence}

We employ two models to evaluate the effect of monetary incentives on dengue
incidence. The first is a differences-in-differences estimation that tests
whether the intervention _per se_ had any effect on dengue hospitalisations in
the city of Rio Verde. The second model compares dengue incidence in each of
the treatment statuses with incidence figures in the control group. The unit of
analysis is the census tract. Table \ref{tab:incidence} displays the results
for our first model, in which we measure the impact of our intervention after 4,
8, 12, and 16 weeks. The last column shows how collective incentives compare
with individual bonuses, with the latter considered as the control group.

\vspace{.3cm}

```{r, echo = FALSE, results='asis', cache=T}
mod <- list()
mod[[1]] <- felm(ncas4weekwindow ~ Tm2+D+DID|0|0|CensusSectors, data = did_full)
mod[[2]] <- felm(ncas8weekwindow ~ Tm2+D+DID|0|0|CensusSectors, data = did_full)
mod[[3]] <- felm(ncas12weekwindow ~ Tm2+D+DID|0|0|CensusSectors, data = did_full)
mod[[4]] <- felm(ncas ~ Tm2+D+DID|0|0|CensusSectors, data = did_full)
stargazer(mod[[1]], mod[[2]], mod[[3]], mod[[4]],
          dep.var.labels = c('\\shortstack{Four \\\\ Weeks Window}',
                             '\\shortstack{Eight \\\\ Weeks Window}',
                             '\\shortstack{Twelve \\\\ Weeks Window}',
                             '\\shortstack{Sixteen \\\\ Weeks Window}'),
          keep.stat = c('ser', 'n'),
          keep = c('DID'),
          notes = c('Robust standard errors clustered at the census tract level in parentheses.'),
          label = 'tab:incidence',
          covariate.labels = c('Dif-in-Dif'),
          dep.var.caption = "Disease Incidence Before and After Intervention:",
          title = 'Differences-in-Differences Model', header = FALSE)
```

The intervention lowered dengue incidence in 10.3% (CI: [-18.9%, -1.6%]) if we take
both treatments together and compare them with the control group. The result is
consistent around fewer weeks before and after the intervention, but it does
not reach conventional levels of statistical significance. This suggests that
the treatment effect is probably small.

\vspace{.3cm}

```{r, echo=FALSE, results='asis', cache=T}
## Health outcome
moutc <- list()
moutc[[1]] <- lm(nc15dafter ~ D1,
                 data = dfr2, weights = dfr2$weightsD1)
moutc[[2]] <- lm(nc30dafter ~ D1,
                 data = dfr2, weights = dfr2$weightsD1)
moutc[[3]] <- lm(nc60dafter ~ D1,
                 data = dfr2, weights = dfr2$weightsD1)
moutc[[4]] <- lm(nc90dafter ~ D1,
                 data = dfr2, weights = dfr2$weightsD1)
moutc[[5]] <- lm(nc30to60dafter ~ D1,
                 data = dfr2, weights = dfr2$weightsD1)
moutc[[6]] <- lm(plac ~ D1,
                 data = dfr2, weights = dfr2$weightsD1)

stargazer(moutc[[6]], moutc[[1]], moutc[[2]],
          moutc[[3]], moutc[[4]], moutc[[5]],
          se = starprep(moutc[[6]], moutc[[1]],
                        moutc[[2]], moutc[[3]],
                        moutc[[4]], moutc[[5]],
                        se_type = "HC1"),
          dep.var.labels = c('\\shortstack{Before \\\\ Int.}',
                             '\\shortstack{Fifteen \\\\ Days \\\\ After Int.}',
                             '\\shortstack{Thirty \\\\ Days \\\\ After Int.}',
                             '\\shortstack{Sixty \\\\ Days \\\\ After Int.}',
                             '\\shortstack{Ninety \\\\ Days \\\\ After Int.}',
                             '\\shortstack{Btw. Thirty \\\\ \\& Sixty Days \\\\ After Int.}'), label = 'tab:census',
          keep.stat = c('ser', 'n'),
          keep = c('D1Individual', 'D1Collective'),
          covariate.labels = c('Individual Bonus',
                               'Collective Bonus'),
          dep.var.caption = "Number of Reported Hospital Cases:",
          title = 'Disease Incidence', header = FALSE,
          font.size = 'small')
```

In our second model, we measure the effect of reward schemes on the number of
hospitalisations per census tract. Table \ref{tab:census} shows hospitalisation
counts before the intervention, and then 15, 30, 60, 90, and between 30 to 60
days after the treatment. Note that 96 pairs of healthcare workers analysed 139
census tracts. This is because 43 of those pairs worked in more than one census
tract at once.

We see that neither treatment has an impact on disease incidence in the Rio
Verde census tracts. The treatment effect is not statistically different from
zero in any model specification. Yet, the coefficients are negative, which
indicates that had an impact occurred, the intervention would have lowered
dengue incidence in the treated areas.

# Discussion
\label{sec:discussion}

We show that monetary incentives increase the productivity of community health
care workers in Brazil. We contrast a control group (which received a flat
performance fee) with two treatment groups, one which received an individual
performance bonus, and second one that received a collective performance
incentive. We find that both treatments increase the number of breeding sites
encountered, and that collective incentives outperform individual performance
rewards when the task is more laborious and excludable (exterminating larvae).
Lastly, the control group presented a high incidence of houses visited in less
than two minutes apart from each other, which suggests that the workers may
not carry their tasks as intended in the absence of incentives. In a nutshell,
monetary incentives work.

Our conclusions are consistent with results from comparable studies conducted
in other settings, such as in education [@duflo2012incentives] and in private
firms [@bandiera2013team]. The findings also lend further evidence that
incentivising teams is particularly effective when they carry out
labour-intensive tasks, which opens a new avenue for research in preventive
healthcare.

Regarding the effect of incentives on city-level dengue incidence, we find only
mixed results. Why did the treatment not work as we planned? We believe the
treatment was not as strong as it needed to be. Our treatment probably did not
cover enough areas to ensure there was no mosquitos around the assigned blocks.
As mosquitos can travel for long distances, the treatment might not have been
sufficient. Also, Rio Verde had significant rates of dengue incidence during
our experiment, reaching 3,411 reported infections from January to September
2018 [@goias2020]. Therefore, the treatment might have stronger effects only
when applied for longer periods of time or in more neighbourhoods.

In this respect, our study suggests is that even though healthcare workers may
be motivated, their efforts do not necessarily lead to a reduction in mosquito
infections. Higher field productivity does not guarantee that the *Aedes
aegypti* mosquito can be eradicated. To our knowledge, this issue has received
little attention in the literature. Future research may explore the reasons why
this is the case and if the disconnect between the two can also be identified
in other areas.

## Limitations
\label{sub:limitations}

The study has several limitations. First, we ran our experiment for a small
period of time and in a particular Brazilian city, thus the findings may not
generalise to other locations. Second, our experiment needs a broader coverage
to measure whether the treatment has a significant impact on dengue incidence,
as the effect size is smaller than we expected. Third, we included a
post-treatment adjustment to ensure the balance of our territorial variables.
Although using matching estimators to improve balance is good statistical
practice, the method requires additional assumptions. Finally, the experiment
only measures the impact of financial rewards on worker performance, and it
does not assess the effect of non-monetary incentives, such as motivational
messages or constant feedback, which may be useful alternatives in
jurisdictions where legal or financial constraints prevent the government from
distributing financial bonuses to civil servants.

## Public Health Implications
\label{sub:implications}

The experiment has important lessons for policy-making. *A. aegypti*-borne
diseases are a significant public health concern, especially in developing
nations. Results of this study indicate that monetary rewards play a major role
in motivating civil servants, and that peer incentives are more cost-effective
than individual bonuses. Finally, the article shows that anti-mosquito
campaigns require a sustained effort to guarantee their success, and we still
have no robust evidence that worker productivity alone is sufficient to reduce
dengue infections.

\newpage
\setlength{\parindent}{0cm}
\setlength{\parskip}{5pt}